Conservative Model Order Reduction of Fluid Flow

### 2018-07-13 sections/5.res.tex
Upload corrections of Incompressible results

\subsection{Vortex Merging}
Consider the $2D$ incompressible Euler equation, \eqref{eq:3.8} with no viscous term, on a square domain $\Omega = [0,2\pi]^2$, provided with periodic boundary conditions. Spatial derivatives are discretized using a Fourier spectral method, which in centered differential operators in the Fourier space. To capture the fine details characterizing the solution, $256\times 256$ modes have been adopted. The initial condition is given in terms of the vorticity distribution $\mathbf{\omega} = \nabla \times u$ {\commnet \{ check this \}}.

We consider the evolution of three different vortices, and the initial structure is given by
\begin{equation}\label{eqn:initial_cond_vort}
\omega = \omega_0 + \sum_{i=1}^{3} \alpha_i e^{-\dfrac{\left(x-x_i\right)^{2}+\left(y-y_i\right)^{2}}{\beta^2}}.
\end{equation}
In \eqref{eqn:initial_cond_vort}, $(x,y)$ represents the spatial coordinates, $\left( x_i, y_i\right)$ is the center of the $i$th vortex, $\alpha_i$ its maximum amplitude, and $\beta$ controls the effective radius of the vortex. In this example, the center of three vortices are positioned at 
\begin{equation}
\left( x_1, y_1 \right) = \left(0.75\pi,\pi\right) , \left( x_2, y_2 \right) = \left(1.25\pi,\pi\right), \left( x_3, y_3 \right) = \left(1.25\pi,1.5\pi\right),
\end{equation}
close to the center of the domain, of which, two have positive spins with $\alpha_1 = \alpha_2 = \pi$ and the other rotates in the opposite direction with $\alpha_3 = -0.5\pi$. The effective radius of all the vortices is set to $\beta = 1 / \pi$. This arrangement of vortices is an effective initial condition for the process of merging vortices. This phenomenon is often a result of fast-moving dipoles {\commnet \{ what is a dipole?\}} facing another vortex {\commnet \{ ref \}}. The forces developed in the merging process transfers the vorticity from the initial configuration, into long, narrow, and spiral-shaped strips of intense vorticity \cite{filaments_vort}. The formation of such thin vorticity filaments in the fluid, can pose numerical challenges.

@@

To define the initial conditions in terms of the velocity components $\omega$ and the pressure $p$, we define a stream-function $\Psi$, the solution to the equation 
\begin{equation} \label{eq:5.1}
	\nabla \Psi = \omega.
\end{equation}
The initial velocity is then given by $\nabla \times \Psi$. To solve the stream-function problem \eqref{eq:5.1}, we require $\int_{\Omega} \omega \ dx= 0$ {\commnet \{ check this \}}. It is easily verified that this requirement implies $\omega_0 = 0.038$. The pressure is recovered by solving the related Poisson pressure equation {\commnet \{ write the equation here \}}. The implicit midpoint scheme, to mimic the time integration scheme presented in \eqref{eq:3.27}, is considered to integrate in time. The merging phenomenon is simulated for a total of $18$ time units using a temporal step $\Delta t=0.004$.

\Cref{fig:unstable_full_models} illustrates the evolution of the kinetic energy for the advective, divergence, and the skew-symmetric form of the high-fidelity system. It is observed that only the skew-symmetric form preserves the kinetic energy, confirming the results in \Cref{sec:skew.2}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Comparison Skew symmetric, divergence and advective form %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \centering
  \begin{tikzpicture}[scale=0.5452]
    \begin{axis}[ylabel=$E_K(t)$,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={Skew Symmetric form, Divergent form, Advective form},
                 legend style={font=\large},
                 width=0.7\linewidth, 
                 height=0.5\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=4},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_skew_symmetric.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_divergence.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_convective.txt};
    \end{axis}%  
  \end{tikzpicture}
  \caption{Conservation of the kinetic energy $K$ for the advective, divergence and the skew-symmetric formulations. {\commnet \{ please change Divergent to Divergence in the legend \}} }
  \label{fig:unstable_full_models}
\end{figure}
A total of $5000$ temporal snapshots is used to construct a reduced basis, following the process discussed in \Cref{sec:mor}. The decay of the singular values, as an indication of the reducibility of the problem, is presented in \Cref{fig:RIC}. The first $35$ POD modes corresponds to over $99\%$ of the {\commnet \{ kinetic? \}} energy of the high fidelity solution. This suggests that an accurate reduced system can be constructed using a small number of basis vectors. To illustrate the effectiveness of the method, smaller bases are also considered.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% Singular values decay merging vorteces %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.5452]
    \begin{semilogyaxis}[ylabel = $\frac{\sigma_j}{\sum_i \sigma_i}$,
                 xlabel=Basis ,
                 label style={font=\large},
                 %legend pos=north east,
                 %legend entries={$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=0.7\linewidth, 
                 height=0.5\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=0,
                 ymax = 2,
                 samples = 100]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = x, y = s] {./data/Incompressible_Euler/Singular_values/sv.txt};
    \end{semilogyaxis}
\end{tikzpicture}
\caption{The decay of the singular values for the merging vortices problem. {\commnet \{please put this figure and figure 1 into a single figure (Left and right to each other)\} } }
\label{fig:RIC}
\end{figure}
\Cref{fig:approx_error_incompressible_Euler} shows the 2-norm error between the high-fidelity solution and the approximated solution. It is seen that the error decreases, consistently, as the number of basis vectors increase. Furthermore, the accuracy is maintained over the period of time integration.

@@

    \begin{semilogyaxis}[ylabel = $\left | \mathbf{v}(t)-\mathbf{v}^r(t) \right |^2$,

@@

    \begin{axis}[ylabel = $|K(t)-K_r^{r}(t)|$,

@@

\caption{(\protect\subref{fig:approx_error_incompressible_Euler}) Evolution of 2-norm error in velocity, between the high-fidelity system and the reduced system. (\protect\subref{fig:energy_error_incompressible_Euler}) Conservation of the kinetic energy. {\commnet \{ in fig a $y$ axis: $\| u - \tilde u \|_2$ \} \{ in fig b $|k - \tilde K|$ \}}  }

>>>>

\subsection{Vortex Merging}
Consider the $2D$ incompressible Euler equation, \eqref{eq:3.8} with no viscous term, on a square domain $\Omega = [0,2\pi]^2$, provided with periodic boundary conditions. Spatial derivatives are discretized using a Fourier spectral method, which in centered differential operators in the Fourier space. To capture the fine details characterizing the solution, $256\times 256$ modes have been adopted. The initial condition is given in terms of the vorticity distribution $\mathbf{\omega} = \nabla \times u$.

We consider the evolution of three different vortices, and the initial structure is given by
\begin{equation}\label{eqn:initial_cond_vort}
\omega = \omega_0 + \sum_{i=1}^{3} \alpha_i e^{-\dfrac{\left(x-x_i\right)^{2}+\left(y-y_i\right)^{2}}{\beta^2}}.
\end{equation}
In \eqref{eqn:initial_cond_vort}, $(x,y)$ represents the spatial coordinates, $\left( x_i, y_i\right)$ is the center of the $i$th vortex, $\alpha_i$ its maximum amplitude, and $\beta$ controls the effective radius of the vortex. In this example, the center of three vortices are positioned at 
\begin{equation}
\left( x_1, y_1 \right) = \left(0.75\pi,\pi\right) , \left( x_2, y_2 \right) = \left(1.25\pi,\pi\right), \left( x_3, y_3 \right) = \left(1.25\pi,1.5\pi\right),
\end{equation}
close to the center of the domain, of which, two have positive spins with $\alpha_1 = \alpha_2 = \pi$ and the other rotates in the opposite direction with $\alpha_3 = -0.5\pi$. The effective radius of all the vortices is set to $\beta = 1 / \pi$. This arrangement of vortices is an effective initial condition for the process of merging vortices. This phenomenon is often a result of fast-moving dipoles of vortices with the same spin facing another vortex \cite{filaments_vort2} of opposite spin. The forces developed in the merging process transfers the vorticity from the initial configuration, into long, narrow, and spiral-shaped strips of intense vorticity \cite{filaments_vort}. The formation of such thin vorticity filaments in the fluid, can pose numerical challenges.

@@

To define the initial conditions in terms of the velocity components $\omega$ and the pressure $p$, we define a stream-function $\Psi$, the solution to the equation 
\begin{equation} \label{eq:5.1}
	\nabla \Psi = \omega.
\end{equation}
The initial velocity is then given by $\nabla \times \Psi$. To solve the stream-function problem \eqref{eq:5.1}, we require $\int_{\Omega} \omega \ dx= 0$. It is easily verified that this requirement implies $\omega_0 = 0.038$. The pressure is recovered by solving the related Poisson pressure equation 
\begin{equation*}
\Delta p = - \nabla \cdot S_{u}(u),
\end{equation*}
obtained by applying the divergence operator to \eqref{eq:3.8} and using the incompressibility condition.
 The implicit midpoint scheme, to mimic the time integration scheme presented in \eqref{eq:3.27}, is considered to integrate in time. The merging phenomenon is simulated for a total of $18$ time units using a temporal step $\Delta t=0.004$.

\Cref{compare_different_models} illustrates the evolution of the kinetic energy for the advective, divergence, and the skew-symmetric form of the high-fidelity system. It is observed that only the skew-symmetric form preserves the kinetic energy, confirming the results in \Cref{sec:skew.2}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Comparison Skew symmetric, divergence and advective form %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \centering
  \begin{subfigure}[]{0.47\linewidth}
  \begin{tikzpicture}[scale=0.58]
    \begin{axis}[ylabel=$E_K(t)$,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={Skew Symmetric form, Divergence form, Advective form},
                 legend style={font=\large},
                 width=1.4\linewidth, 
                 height=1\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=4},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_skew_symmetric.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_divergence.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_convective.txt};
    \end{axis}%  
  \end{tikzpicture}
\caption{}
\label{compare_different_models}
\end{subfigure}
  \begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{semilogyaxis}[ylabel = $\frac{\sigma_j}{\sum_i \sigma_i}$,
                 xlabel=Basis POD,
                 label style={font=\large},
                 %legend pos=north east,
                 %legend entries={$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=0,
                 ymax = 2,
                 samples = 100]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = x, y = s] {./data/Incompressible_Euler/Singular_values/sv.txt};
    \end{semilogyaxis}
\end{tikzpicture}
\caption{}
\label{singular_values_decay}
\end{subfigure}
  \caption{(\protect\subref{compare_different_models}) Conservation of the kinetic energy $K$ for the advective, divergence and the skew-symmetric formulations. (\protect\subref{singular_values_decay}) The decay of the singular values for the merging vortices problem. }
  \label{fig:unstable_full_models}
\end{figure}
A total of $5000$ temporal snapshots is used to construct a reduced basis, following the process discussed in \Cref{sec:mor}. The decay of the singular values, as an indication of the reducibility of the problem, is presented in \Cref{singular_values_decay}. The first $35$ POD modes corresponds to over $99\%$ of the modes energy of the high fidelity solution. This suggests that an accurate reduced system can be constructed using a small number of basis vectors. To illustrate the effectiveness of the method, smaller bases are also considered.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% Singular values decay merging vorteces %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Cref{fig:approx_error_incompressible_Euler} shows the 2-norm error between the high-fidelity solution and the approximated solution. It is seen that the error decreases, consistently, as the number of basis vectors increase. Furthermore, the accuracy is maintained over the period of time integration.

@@

    \begin{semilogyaxis}[ylabel = $\| \mathbf{u}-\tilde{\mathbf{u}}\|_2$,

@@

    \begin{axis}[ylabel = $|K-\tilde{K}|$,

@@

\caption{(\protect\subref{fig:approx_error_incompressible_Euler}) Evolution of 2-norm error in velocity, between the high-fidelity system and the reduced system. (\protect\subref{fig:energy_error_incompressible_Euler}) Conservation of the kinetic energy.}
